---
title: "Gaussian test cases"
output: html_notebook
---

```{r}
source("augmented_utils.R")
```


```{r}
registerDoParallel(cores = 20)


pp = transport::pp
```


```{r}
CtoR = function(clusts, only_bornes = TRUE, n_sample = NULL){
  return(list(try_norm(clusts, only_bornes, n_sample), as.list(rep("normal", length(clusts)))))
  }
```

We build a function that perform the GMM algorithm
```{r}
GMM = function(samp, nb_start=1){
  list_start = list() 
  for(start in 1:nb_start){
    m = sample(samp, 2) #We choose initial means
    list_start[[start]] = m 
    #print(m)
    sigma = std(samp) 
    variances = rep(sigma^2/2,2) #The initial variances are related to the variance of the sample
    pi_i = rep(1,2) #the initial weights are equal
    vec_res = c(m, variances, pi_i/sum(pi_i)) 
    vec_res_new = rep(10^5, length(vec_res))
    best_diss = 10^5
    converge = c()
    it = 0
    while(sum(abs(vec_res_new-vec_res)) > 10^-6 & it<10^3){
      it = it+1
      vec_res_new = vec_res
      ##Expectation
      p = Vectorize(function(i){Vectorize(function(j){pi_i[j]*dnorm(samp[i], m[j], sqrt(variances[j]))})(1:2)})(1:length(samp)) #likelihood of each element
      s = apply(p, 2, sum)
      p = t(t(p)/s) #normalized likelihoods

    pi_i = apply(p, 1, sum) #new weights 
    
    #Maximisation
    m = Vectorize(function(i){weighted.mean(samp, p[i,])})(1:2) #means
    variances = Vectorize(function(i){weighted.mean((samp-m[i])^2, p[i,])})(1:2) #variances

  vec_res= c(m, variances, pi_i/sum(pi_i))
  converge = c(converge,sum(abs(vec_res_new-vec_res)))
  converge_bool = sum(sort(tail(converge, 10)) == tail(converge,10)) != 10
    }
  mixture = c(rnorm(round(pi_i[1]/sum(pi_i),5)*10^5, m[1], sqrt(variances[1])),  rnorm(round(pi_i[2]/sum(pi_i),5)*10^5, m[2], sqrt(variances[2])))
  diss = wasserstein1d(mixture, samp, p=2)
  if(diss < best_diss){ #keep the best start regarding the global error
    best_params = vec_res
  
  }
    
  }
  return(list(best_params[5:6],best_params[1:2], sqrt(best_params[3:4]), list_start))
}

```

## Testing samples

Generation of 15 test samples

```{r}
RNGkind("default")
set.seed(123)
list_sampgauss = list()
df_params_gauss = data.frame()
for(i in 1:15){
  params = runif(4)
  weights = runif(1,0.3,0.7)
  weights = c(weights, 1-weights)
  lengths = round(weights*400)
  list_sampgauss[[i]] = matrix(c(sobol(10000, normal = TRUE)[(10000-lengths[1]+1):10000]*params[2]+params[1], sobol(2000, normal = TRUE)[(2000-lengths[2]+1):2000]*params[4]+params[3]))
  df_params_gauss = rbind(df_params_gauss, c(params, lengths))
}
```

GMM and augmented quantization for the test samples

```{r}
set.seed(5)
list_record_mixtures_gauss = list()
list_best_mixtures_gmm = list()
list_error_plot = list()
for(i in 1:15){
  set.seed(i,"L'Ecuyer-CMRG")
  print(paste("SAMPLE NUMERO ",i))
  res_gmm = GMM(list_sampgauss[[i]], nb_start = 3) #Perform GMM
  list_best_mixtures_gmm[[i]] = res_gmm
  mixt_gmm = get_mixture(rep = list(matrix(c(res_gmm[[2]][1],res_gmm[[3]][1])), matrix(c(res_gmm[[2]][2],res_gmm[[3]][2]))), weights = res_gmm[[1]], law = list("normal","normal"), n=10^6)
  dist_gmm = wasserstein1d(mixt_gmm, list_sampgauss[[i]],p=2) #Compute global error
  print(paste("dist gmm", dist_gmm))
  print("GMM Done")
  sigma = std(list_sampgauss[[i]])/sqrt(2)
  err = 10^5
  for(start in 1:3){
    m = res_gmm[[4]][[start]] #get the parameters of the starts for GMM
    rep = list(matrix(c(m[1],sigma)), matrix(c(m[2],sigma)))
    res_gauss_1d = augmented_quanti(list_sampgauss[[i]], rep, law = list("normal","normal"), n_sample = 3000, l_bin =2,vec_prop = c(0.4,0.2,0.1),it_lim = 5,find_c_bool = TRUE ) #perform Augmented quanti
    best = find_best(res_gauss_1d)
    for(j in 1:length(best[[4]][[1]])){
      if(best[[4]][[1]][[j]][2,1]<=0){best[[4]][[1]][[j]][2,1] = 10^(-15)}
    }
    mixt_augm = get_mixture(rep = best[[4]][[1]], clusts = best[[3]],law = list("normal","normal"),n = 10^6)

    dist_augm = wasserstein1d(mixt_augm, list_sampgauss[[i]],p=2)
    print(paste("dist augm", dist_augm))

    if(dist_augm<err){
      list_record_mixtures_gauss[[i]] = res_gauss_1d
      err = dist_augm
    }
    
  }

}




```

Compute the quantization and global errors

```{r}
RNGkind("default")
set.seed(10)

error_aug = c()
wass_aug = c()
wass_true = c()
error_true = c()
wass_gmm = c()

for(i in 1:length(list_record_mixtures_gauss)){
  print(i)
  a = as.numeric(df_params_gauss[i,1])
  b=as.numeric(df_params_gauss[i,2])
  c = as.numeric(df_params_gauss[i,3])
  d = as.numeric(df_params_gauss[i,4])
  
  best = find_best(list_record_mixtures_gauss[[i]])
      for(j in 1:length(best[[4]][[1]])){
      if(best[[4]][[1]][[j]][2,1]<=0){best[[4]][[1]][[j]][2,1] = 10^(-15)}
    }
  clusts = best[[3]]
  weights = Vectorize(function(j){nrow(clusts[[j]])/sum(sapply(clusts, nrow))})(1:length(clusts))
  reps = lapply(1:2,function(k){matrix(rnorm(10^6, best[[4]][[1]][[k]][1,1], best[[4]][[1]][[k]][2,1]))})
  error_aug = c(error_aug,sqrt(weighted.mean(local_errors(clusts, reps, 1:2)^2, weights)))
  
  mixture = get_mixture(rep = best[[4]][[1]], clusts = best[[3]],law = list("normal","normal"),n = 10^6)
  wass_aug = c(wass_aug, wasserstein1d(mixture, list_sampgauss[[i]],p=2))
    
  reps_true = list(matrix(rnorm(10^6, a, b)), matrix(rnorm(10^6, c, d)))
  clusts_true = list(matrix(list_sampgauss[[i]][1:df_params_gauss[i,5],]), matrix(list_sampgauss[[i]][(df_params_gauss[i,5]+1):nrow(list_sampgauss[[i]]),]))
  error_true = c(error_true,sqrt(weighted.mean(local_errors(clusts_true, reps_true, 1:2)^2, as.numeric(df_params_gauss[i, 5:6]/nrow(list_sampgauss[[i]]))))) 
  
    mixt_true = get_mixture(rep = list(matrix(as.numeric(df_params_gauss[i, 1:2])), matrix(as.numeric(df_params_gauss[i, 3:4]))), weights = as.numeric(df_params_gauss[i, 5:6]/nrow(list_sampgauss[[i]])), law = list("normal","normal"), n=10^6)
    wass_true = c(wass_true, wasserstein1d(mixt_true, list_sampgauss[[i]], p=2))
  res_gmm = list_best_mixtures_gmm[[i]]
  mixt_gmm = get_mixture(rep = list(matrix(c(res_gmm[[2]][1],res_gmm[[3]][1])), matrix(c(res_gmm[[2]][2],res_gmm[[3]][2]))), weights = res_gmm[[1]], law = list("normal","normal"), n=10^6)
  mixt_true = get_mixture(rep = list(matrix(as.numeric(df_params_gauss[i, 1:2])), matrix(as.numeric(df_params_gauss[i, 3:4]))), weights = as.numeric(df_params_gauss[i, 5:6]/nrow(list_sampgauss[[i]])), law = list("normal","normal"), n=10^6)
  wass_gmm = c(wass_gmm, wasserstein1d(mixt_gmm, list_sampgauss[[i]],p=2))


}

```

Plot the errors

```{r}
df_errors = data.frame(rbind(cbind(wass_true,3),cbind(wass_aug,4), cbind(wass_gmm,5), cbind(error_aug,2), cbind(error_true,1)))
colnames(df_errors) = c("Error","type")
df_errors$type = as.factor(df_errors$type)


ggplot(df_errors[df_errors$type %in% c('1','2'),], aes(x = type, y = Error, col = type)) + geom_boxplot() + theme_bw() + scale_color_discrete(name = "Quantization error", labels = c("True", "Augmented")) + theme(legend.position="top",legend.text = element_text(size=15),legend.title = element_text("size" = 15))

ggplot(df_errors[df_errors$type %in% c('3','4','5'),], aes(x = type, y = Error, col = type)) + geom_boxplot() + theme_bw() + scale_color_manual(name = "Global error", labels = c("True","Augmented","GMM"), values = c("#F8766D","#00B9E3","#00BA38")) + theme(legend.position="top",legend.text = element_text(size=15), legend.title = element_text("size" = 15))

```

Plot the evolution of the errors through the iterations

```{r}
df_plot_err = data.frame()
for(i in 1:15){
  err = get_evol_error(list_record_mixtures_gauss[[i]])
  df_plot_err = rbind(df_plot_err, cbind(1:length(err),err, col = i))
}
df_plot_err$col = as.factor(df_plot_err$col)

ggplot(df_plot_err)+ geom_line(aes(x = V1, y = err,col = col))  +  theme_bw()+ theme(legend.position = "none")+ scale_y_continuous(trans='log10') 
```

## Focus on an example

We select the sample for which the difference between the augmented quantization error and the true quantization error is the median of the 15 obtained.

```{r}

set.seed(10)
vec_diff =  wass_aug-wass_gmm 
idx = which.min(abs(vec_diff-quantile(vec_diff,0.5))) #Select this sample
idx = 13
idx = which.min(abs(vec_diff-mean(vec_diff)))
best = find_best(list_record_mixtures_gauss[[idx]])
weights = df_params_gauss[idx,5:6]/nrow(list_sampgauss[[idx]]) #Get the true weights
rpz1 = as.character(round(df_params_gauss[idx,1:2],2)) 
rpz1[2] = paste(rpz1[2], "²", sep = "")
rpz1 = paste(rpz1, collapse = ",")
rpz1 = paste("N(", rpz1,")", sep = "")
rpz2 = as.character(round(df_params_gauss[idx,3:4],2))
rpz2[2] = paste(rpz2[2], "²", sep = "")
rpz2 = paste(rpz2, collapse = ",")
rpz2 = paste("N(", rpz2,")", sep = "")

plot_sample(list(list_sampgauss[[idx]][1:df_params_gauss[idx,5],], list_sampgauss[[idx]][(df_params_gauss[idx,5]+1):nrow(list_sampgauss[[idx]]),]),rpz = c(rpz1, rpz2), colnames = c("x")) #plot the sample


#plot the densities of the obtained clusters and their associated representatives
plot(density(best[[3]][[1]]), xlim = c(-2,4),ylim = c(0,4.5), xlab = "x")
lines(density(best[[3]][[2]])) 

lines(density(rnorm(10^6,best[[4]][[1]][[1]][1,1], best[[4]][[1]][[1]][2,1])), col = "red",lty = "dashed")
lines(density(rnorm(10^6,best[[4]][[1]][[2]][1,1], best[[4]][[1]][[2]][2,1])), col = "red",lty = "dashed")

legend(x = "topright",
       legend = c("Representatives", "Clusters"),
       lty = c(2, 1),
       col = c("red", "black"))
title(xlab = "x")

mixture = get_mixture(rep = best[[4]][[1]], clusts = best[[3]],law = list("normal","normal"),n = 10^6) #get the obtained mixture

res_gmm = list_best_mixtures_gmm[[idx]]
mixt_gmm = get_mixture(rep = list(matrix(c(res_gmm[[2]][1],res_gmm[[3]][1])), matrix(c(res_gmm[[2]][2],res_gmm[[3]][2]))), weights = res_gmm[[1]], law = list("normal","normal"), n=10^6) #get the gmm mixture

mixt_true = get_mixture(rep = list(matrix(as.numeric(df_params_gauss[idx, 1:2])), matrix(as.numeric(df_params_gauss[idx, 3:4]))), weights = as.numeric(df_params_gauss[idx, 5:6])/nrow(list_sampgauss[[idx]]), law = list("normal","normal"), n=10^6) #get the true mixture


#plot the distributions of these 3 mixtures
plot(NA, ylim = c(0,2), xlab = "x",xlim = c(-2,4), ylab  = "Density")
hist(list_sampgauss[[idx]],freq = FALSE, breaks = 50, col = NULL,add=TRUE)

lines(density(mixture), col = "red")
lines(density(mixt_gmm), col = "green")

legend(x = "topright",     
       legend = c("Sample", "AQ", "GMM"), lty = c(1,1,1),
       col = c("black", "red", "green"))


c(error_true[idx], error_aug[idx], wass_true[idx], wass_aug[idx], wass_gmm[idx])

df_params_gauss[idx,]
best[[4]]
list(matrix(c(res_gmm[[2]][1],res_gmm[[3]][1])), matrix(c(res_gmm[[2]][2],res_gmm[[3]][2])))
```



