---
title: "Hybrid test cases"
output: html_notebook
---

```{r}
source("augmented_utils.R")

```

```{r}
  CtoR = function(clusts, only_bornes = TRUE, n_sample = 500){
 norm = try_norm(clusts, only_bornes = TRUE)
 norm_tilde = from_law_to_sample(norm, law = as.list(rep("normal", length(clusts))), n = n_sample)
 unif = try_unif(clusts, only_bornes = TRUE)
 unif_tilde = from_law_to_sample(unif, law = as.list(rep("unif", length(clusts))), n = n_sample)
  diss_norm = sapply(1:length(clusts), function(j){wasserstein1d(as.numeric(norm_tilde[[j]]), as.numeric(clusts[[j]]), p=2)})
  diss_unif = sapply(1:length(clusts), function(j){wasserstein1d(as.numeric(unif_tilde[[j]]), as.numeric(clusts[[j]]), p=2)})
 diff_diss = diss_norm-diss_unif
 res = lapply(1:length(clusts), function(j){
   if(diff_diss[j] <= 0){
     if(only_bornes){return(list(norm[[j]],"normal"))}
     else{return(list(norm_tilde[[j]], "normal"))}
   }
   else{
      if(only_bornes){return(list(unif[[j]],"unif"))}
     else{return(list(unif_tilde[[j]],"unif"))}
   }
 })
 res = list(lapply(res, function(x){x[[1]]}), lapply(res, function(x){x[[2]]}))
 return(res)
}

```

# Simple test case

We introduce the sample representing the mixture $R_{J}$ with 

- $R_{1} = \mathcal{U}(0.2,0.5)$
- $R_{2} = \mathcal{N}(0.6,0.2^2)$
- $P(J=1) = \frac{100}{350}$
- $P(J=2) = \frac{250}{350}$


```{r}
set.seed(4) 
size_unif = 100
size_gauss = 350-size_unif
pi = base::pi
samp_norm = matrix(c(sobol(size_unif)*0.3+0.2,0.6+ 0.2*sobol(200000,normal = TRUE)[(20000-size_gauss+1):20000]))

plot_sample(list(samp_norm[1:size_unif,], samp_norm[(size_unif+1):nrow(samp_norm),]), colnames = c("x"), rpz = c("U(0.2,0.5)","N(0.6,0.04)"))
```


```{r}
set.seed(10,"L'Ecuyer-CMRG")
rep_init = list()
rep_init[[1]] = matrix(c(0,0.3))
rep_init[[2]] = matrix(c(0.3,1))

law = list("unif","unif")

 res_hybrid = augmented_quanti(samp_norm, rep_init, law = law, n_sample = 3000,l_bin = 2, threshold = 0.01,vec_prop = c(0.4,0.3,0.2,0.1,0.05))
 
best_hybrid = find_best(res_hybrid)

best_hybrid[[4]]
sapply(best_hybrid[[3]], nrow)/length(samp_norm)
sqrt(best_hybrid[[5]])

```

```{r}
plot_sample(list(samp_norm[1:size_unif,], samp_norm[(size_unif+1):nrow(samp_norm),]), colnames = c("x"), rpz = c("U(0.2,0.5)","N(0.6,0.04)"))
plot_clusts(best_hybrid[[3]])



reps_true = list(matrix(runif(10^6, 0.2, 0.5)), matrix(rnorm(10^6, 0.6, 0.2)))

reps_aug = list(matrix(rnorm(10^6, best_hybrid[[4]][[1]][[1]][1,1], best_hybrid[[4]][[1]][[1]][2,1])), matrix(runif(10^6, best_hybrid[[4]][[1]][[2]][1,1],best_hybrid[[4]][[1]][[2]][2,1])))

clusts_true = list(samp_norm[1:size_unif,], samp_norm[(size_unif+1):nrow(samp_norm),])
clusts_aug = best_hybrid[[3]]
qe_true = sqrt(weighted.mean(local_errors(clusts_true, reps_true, 1:2)^2, c(size_unif,nrow(samp_norm)-size_unif)))
qe_aug =sqrt(weighted.mean(local_errors(clusts_aug, reps_aug, 1:2)^2, sapply(clusts_aug,nrow)))
mixture = get_mixture(rep = best_hybrid[[4]][[1]], clusts = best_hybrid[[3]],law = best_hybrid[[4]][[2]],n = 10^6)
ge_aug = wasserstein1d(mixture, samp_norm,p=2)
mixt_true = get_mixture(rep = list(matrix(c(0.2,0.5)), matrix(c(0.6,0.2))), weights = c(size_unif, nrow(samp_norm)-size_unif)/nrow(samp_norm), law = list("unif","normal"), n=10^6)
ge_true = wasserstein1d(mixt_true, samp_norm, p=2)

```

## Tests other samples

We create 15 other samples

```{r}
RNGkind("default")

df_params = data.frame()
list_samp_hybrid = list()
error_true = c()
for(i in 1:15){
  set.seed(i)
  weight = runif(1,0.2,0.5)
  size_unif = as.integer(350*weight) #Size of the uniform sample
  size_norm = 350-size_unif #size of the gaussian sample
  params_unif = sort(runif(2)) #parameters of the uniform distribution
  params_gauss = runif(2) #parameters of the gaussian distribution
  df_params = rbind(df_params, c(params_unif, params_gauss, size_unif, size_norm)) #store the parameters
  pi = base::pi
  list_samp_hybrid[[i]] = matrix(c(sobol(size_unif)*(params_unif[2]-params_unif[1])+params_unif[1],params_gauss[1]+ params_gauss[2]*sobol(2000,normal = TRUE)[(2000-size_norm+1):2000])) #generation of the sample
  reps_true = list(matrix(runif(10^6, params_unif[1], params_unif[2])), matrix(rnorm(10^6, params_gauss[1], params_gauss[2])))
  clusts_true = list(matrix(list_samp_hybrid[[i]][1:size_unif,]), matrix(list_samp_hybrid[[i]][(size_unif+1):length(list_samp_hybrid[[i]]),]))
  error_true = c(error_true,sqrt(weighted.mean(local_errors(clusts_true, reps_true, 1:2)^2, c(weight, 1-weight))) ) #computation of the true computation error
}


```

We perform the Augmented Quantization method on each of the 15 samples

```{r}
list_res = list()
for(i in 1:15){
  err = 10^5
  print(i)
  print(error_true[i])
  for(start in 1:3){
    set.seed(i*3*start,"L'Ecuyer-CMRG")
    rep_init = list()
    rep_init[[1]] = matrix(sort(runif(2)))
    rep_init[[2]] = matrix(sort(runif(2)))
    law = rep("unif",2)
    res_hybrid = augmented_quanti(list_samp_hybrid[[i]], rep_init, law = law, n_sample = 3000,l_bin = 2, threshold = 0.01,vec_prop = c(0.4,0.2,0.1), it_lim = 5) 
    best = find_best(res_hybrid)
    print(sqrt(best[[5]]))
    if(best[[5]]<err){
       list_res[[i]] = res_hybrid
       err = best[[5]]
    }
  }
}

```




We compute the global and quantization errors for each of the sample

```{r}
error_aug = c()
wass_true = c()
wass_aug = c()
for(i in 1:15){
  print(i)
  best = find_best(list_res[[i]])
  
  # reps_true = list(matrix(runif(10^6, df_params[i,1], df_params[i,2])), matrix(rnorm(10^6, df_params[i,3], df_params[i,4])))
  reps_aug = lapply(1:2, function(k){
    if((best[[4]][[2]][[k]] == "unif")){matrix(runif(10^6, best[[4]][[1]][[k]][1,1],best[[4]][[1]][[k]][2,1]))}
    else{matrix(rnorm(10^6, best[[4]][[1]][[k]][1,1],best[[4]][[1]][[k]][2,1]))}})
    
  # clusts_true = list(matrix(list_samp_hybrid[[i]][1:df_params[i,5],]), matrix(list_samp_hybrid[[i]][(df_params[i,5]+1):nrow(list_samp_hybrid[[i]]),]))
  clusts_aug = best[[3]]

  # error_true = c(error_true, sqrt(weighted.mean(local_errors(clusts_true, reps_true, 1:2)^2,sapply(clusts_true,nrow))))
  error_aug = c(error_aug, sqrt(weighted.mean(local_errors(clusts_aug, reps_aug, 1:2)^2, sapply(clusts_aug,nrow))))
  
  mixture = get_mixture(rep = best[[4]][[1]], clusts = best[[3]],law = best[[4]][[2]],n = 10^6)
  wass_aug = c(wass_aug,wasserstein1d(mixture, list_samp_hybrid[[i]],p=2))
  mixt_true = get_mixture(rep = list(matrix(as.numeric(df_params[i, 1:2])), matrix(as.numeric(df_params[i, 3:4]))), weights = as.numeric(df_params[i, 5:6]/nrow(list_samp_hybrid[[i]])), law = list("unif","normal"), n=10^6)
  wass_true = c(wass_true,wasserstein1d(mixt_true, list_samp_hybrid[[i]], p=2))

  
  
  # best = find_best(list_res[[i]])
  # rpz1 = matrix(runif(10^6, df_params[i,1], df_params[i,2]))
  # rpz2 = matrix(rnorm(10^6, df_params[i,3], df_params[i,4]))
  # clusts = list(list_samp_hybrid[[i]][1:df_params[i,5],], list_samp_hybrid[[i]][(df_params[i,5]+1):350,])
  #   error_true = c(error_true, sqrt(weighted.mean(local_errors(clusts, list(rpz1, rpz2), 1:2)^2, c(df_params[i,5]/350, df_params[i,6]/350))))
  #   error_aug = c(error_aug, sqrt(best[[5]]))
  #     mixture = get_mixture(rep = best[[4]][[1]], clusts = best[[3]],law = best[[4]][[2]],n = 10^6)
  # wass_aug = c(wass_aug, wasserstein1d(mixture, list_samp_hybrid[[i]],p=2))
  # 
  #   wass_true = c(wass_true, wasserstein1d(mixt_true, list_samp_hybrid[[i]], p=2))
}

```


We plot the distributions of these errors 


```{r}
df_errors = data.frame(rbind(cbind(wass_true,3),cbind(wass_aug,4), cbind(error_aug,2), cbind(error_true,1)))
colnames(df_errors) = c("Error","type")
df_errors$type = as.factor(df_errors$type)


ggplot(df_errors[df_errors$type %in% c('1','2'),], aes(x = type, y = Error, col = type)) + geom_boxplot() + theme_bw() + scale_color_discrete(name = "Quantization error", labels = c("True", "Augmented"))+ theme(legend.position="top",legend.text = element_text(size=15), legend.title = element_text("size" = 15))

ggplot(df_errors[df_errors$type %in% c('3','4'),], aes(x = type, y = Error, col = type)) + geom_boxplot() + theme_bw() + scale_color_discrete(name = "Global error", labels = c("True","Augmented")) + theme(legend.position="top",legend.text = element_text(size=15), legend.title = element_text("size" = 15))
```


We plot the evolution of the errors through the iterations


```{r}
df_plot_err = data.frame()
for(i in 1:15){
  err = get_evol_error(list_res[[i]])
  df_plot_err = rbind(df_plot_err, cbind(1:length(err),get_evol_error(list_res[[i]]), col = i))
}
df_plot_err$col = as.factor(df_plot_err$col)

ggplot(df_plot_err)+ geom_line(aes(x = V1, y = V2,col = col))  +   theme_bw()+ theme(legend.position = "none") + scale_y_continuous(trans='log10')



```


```
augmented_quanti_hybrid = function(samp, rep, law = "unif", n_sample = 500, l_bin =2, vec_prop = c(0.4,0.3,0.2,0.1,0.05,0.025),it_lim = 10, nb_start = 10, threshold = 0.01){
  rep_bis = from_law_to_sample(rep = rep, law = law, n = n_sample )
  record = list()
  for(pbin in vec_prop){
    record[[paste(pbin)]] = as.list(rep(0,it_lim))
    rep_old = rep
    for(it in 1:it_lim){
      print(it)
      print(rep)
      print(law)
      res_find_clusts = find_clusters_multi_start(samp, rep_bis, nb_start = nb_start)
      clusts = lapply(res_find_clusts[[1]], function(x){matrix(x,ncol = ncol(rep[[1]]))})
      nums = res_find_clusts[[2]]
      if(0 %in% sapply(clusts, nrow)){
        for(i0 in which(sapply(clusts, nrow)== 0)){
          idx_min = which.min(Vectorize(function(i){dist2(samp[i,],rep_bis[[i0]])})(1:nrow(samp)))
          nums[[i0]] = idx_min

          nums[setdiff(1:length(clusts),i0)] = lapply(setdiff(1:length(clusts),i0), function(x){nums[[x]][nums[[x]] != idx_min]})
        }
        clusts = lapply(nums, function(x){matrix(samp[x,], ncol = ncol(samp))})
      }


      errors = local_errors(clusts, CtoR(clusts = clusts, only_bornes = FALSE, n_sample = n_sample)[[1]], idx = 1:length(clusts))
      former_clusts = clusts
      index_throw = order(errors, decreasing = T)[1:l_bin]

      num_throw = lapply(index_throw, function(i){f_delete(clusts[[i]], n = pbin*nrow(clusts[[i]]), n_sample = n_sample)})

      clusts_throw = lapply(1:l_bin, function(i){matrix(clusts[[index_throw[[i]]]][num_throw[[i]],], ncol = ncol(samp))})
      clusts_split = lapply(1:l_bin, function(i){matrix(clusts[[index_throw[[i]]]][-num_throw[[i]],], ncol = ncol(samp))})
      if(length(errors) > l_bin){
        clusts_split = c(clusts_split, lapply(setdiff(1:length(clusts), index_throw), function(i){clusts[[i]]}))
      }
      clusts_split = c(clusts_split, clusts_throw)
      merge = merge_clusts(clusts_split, length(rep), n_sample = n_sample)
      clusts = merge[[2]]
      rep = merge[[1]][[1]]
      law = merge[[1]][[2]]
      error = merge[[4]]
      #print(rep)
      record[[paste(pbin)]][[it]] = list(former_clusts, clusts_split, clusts, list(rep, law), error)
      record<<- record
      for(i in 1:length(rep)){
        if(sum(is.na(rep[[i]])) == 4){
          set.seed(pbin*10^3+it)
          rep[[i]] = matrix(runif(length(rep[[1]])), ncol=2)
    }
  }
      rep_bis = from_law_to_sample(rep, law,n_sample)
      if(sum(sapply(rep, length) == sapply(rep_old, length)) == length(rep)){
      if(sum(Vectorize(function(m){sum(abs(rep[[m]]- rep_old[[m]]))})(1:length(rep))) < threshold){break}
      }
    rep_old = rep
    }
  }
  return(record)
}
```



